---
title: Transcriptomics of _Picea rubens_
author: \emph{Thomas O'Leary}
output:
  pdf_document:
#    latex_engine: xelatex
    keep_tex: true
header-includes:
  - \usepackage[cmintegrals,cmbraces]{newtxmath}
  - \usepackage{ebgaramond}
  - \usepackage[T1]{fontenc}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{wrapfig}
  - \usepackage[font={footnotesize}]{caption}
  # - \usepackage{fancyhdr}
  # - \pagestyle{fancy}
  # - \fancyhead[RO,RE]{O'Leary}
bibliography: tso_refs.bib
csl: mla.csl
fontsize: 12pt
geometry: "left=1cm,right=1cm,top=1cm,bottom=1cm"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
require(tidyverse)
require(kableExtra)
require(DESeq2)
require(scales)
require(ggpubr)
require(wesanderson)
require(vsn)
require(pheatmap)

# Set your working directory
# setwd(here::here("myresults/transcript_counts"))
# source(here::here("load_transcript_data.R"))
```

# Background

### Transcriptomic multiple stressors experimental design

Some _Picea rubens_ individuals, a red spruce that thrives in cool and moist climates, are increasingly experiencing periods of environmental stress, including excessive heat and drought conditions in some of the areas in its range [CITATION??]. This may be due to microclimate differences, such as increased sun exposure and drier soil, but may also be due to larger scale changes in regional climate [CITATION???]. Ten maternal families from two different source climates (hot and dry climate or cool and wet climate) were used. The source climate was determined based on historical climate data on the amount of precipitation in the warmest quarter and the temperature during the hottest month. Individuals from these maternal families were exposed to three experimental treatments: a control, heat stress, and combined heat and drought stress. Each of these treatments included a 16:8 light to dark photoperiod, with variable day and night time temperatures. The control treatment involved 

The purose of this study is to decribe the transcriptomic response of _P. rubens_ under temperature and drought stress




### Sample collection, library preparation, and sequencing

At each collection time point during the course of the experiment, RNA was extracted from whole seedlings (_i.e._ root, stem, and needle tissue). After extraction, RNA quantity and quality was assessed on a Bioanalyzer. 

Samples (RNA concentration > 1 ng/$\mu$L) were sent to Cornell for library preparation and sequencing. Library preparation followed the Lexogen QuantSeq 3’ mRNA-Seq protocols. The samples were sequenced in 86 bp reads on a NextSeq500.
	
# Bioinformatics Pipeline

The quality of the raw reads were accessed using FastQC v$0.11.3$ [@andrews2012]. Low quality sequence data was trimmed from the raw reads using Trimmomatic v$0.33$ [@Bolger2014] and the cleaned reads were vizualized again with FastQC. Those cleaned reads were then mapped to a Norway spruce, _Picea abies_, reference transcriptome [CONGENIE>ORG CITATION] using salmon v$1.1.0$ [@Patro2017]. The reference transcriptome included all coding sequence as well as two kilobases downstream to account for the 3' tag sequencing technique.

Count files were imported into R v$3.6.1$ [@rcore] with the package tximport v$1.12.3$ [@tximport]. Transcripts with less than one transcript per sample were pre-filtered from the analysis. Differential gene expression was called and an FDR of 0.05 using DESeq2 v$1.24.0$ [@Love2014]. Further analysis and visualization was done using the packages ggplot2 and dplyr v$1.2.1$ [@tidyverse].

# Results

```{r, fig.width = 3, fig.height = 2, fig.align = "left"}
  # kable("latex", booktabs = T,
  #       align = c('l', 'c', 'c', 'c'),
  #       caption = "CAPTION") %>%
  # kable_styling(font_size = 10)
```

# ```{r, fig.width = 5, fig.height = 2, fig.cap = "CAPTION")}
# 
# # cowplot::plot_grid(p1, p2, nrow = 1, labels = "AUTO", label_size = 10)
# ```

The mapping rates...

The number of DEGs...

The experimental model ...

Which DEGs...

# Conclusion

The conservation of _Picea rubens_ along the eastern United States...

Future directions for this study must involve...

\pagebreak

# References [^1]

[^1]: This does not exactly follow the MLA citation example you gave, but it was the best `.csl` I could find in time.