---
title: Transcriptomics of _Picea rubens_
author: \emph{Thomas O'Leary}
output:
  pdf_document:
#    latex_engine: xelatex
    keep_tex: true
header-includes:
  - \usepackage[cmintegrals,cmbraces]{newtxmath}
  - \usepackage{ebgaramond}
  - \usepackage[T1]{fontenc}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{wrapfig}
  - \usepackage[font={footnotesize}]{caption}
  # - \usepackage{fancyhdr}
  # - \pagestyle{fancy}
  # - \fancyhead[RO,RE]{O'Leary}
bibliography: tso_refs.bib
csl: mla.csl
fontsize: 12pt
geometry: "left=1cm,right=1cm,top=1cm,bottom=1cm"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
require(tidyverse)
require(kableExtra)
require(DESeq2)
require(scales)
require(ggpubr)
require(wesanderson)
require(vsn)
require(pheatmap)

# Set your working directory
setwd(here::here("myresults/transcript_counts"))



```

# Background

### Transcriptomic multiple stressors experimental design

Some _Picea rubens_ individuals, a red spruce that thrives in cool and moist climates, are increasingly experiencing periods of environmental stress, including excessive heat and drought conditions in some of the areas in its range. [CITATION??] This may be due to microclimate differences, such as increased sun exposure and drier soil. But may also be due to 




The purose of this study is to decribe the transcriptomic response of _P. rubens_ under these environmental stressors.


### Sample collection, library preparation, and sequencing

At each collection time point during the course of the experiment, RNA was extracted from whole seedlings (_i.e._ root, stem, and needle tissue). After extraction, RNA quantity and quality was assessed on a Bioanalyzer. 

Samples (RNA concentration > 1 ng/$\mu$L) were sent to Cornell for library preparation and sequencing. Library preparation followed the Lexogen QuantSeq 3’ mRNA-Seq protocols. The samples were sequenced in 86 bp reads on a NextSeq500.
	
# Bioinformatics Pipeline

The quality of the raw reads were accessed using FastQC v$0.11.3$ [@andrews2012]. To remove low quality sequence data the raw fastq files were trimmed using Trimmomatic v$0.33$ [@Bolger2014] and the cleaned reads were vizualized again with FastQC. The cleaned reads were mapped to a reduced verion of the Norway spruce _Picea abies_ reference genome [@Nystedt2013] using bwa v$0.7.12-r1039$ [@Li2009]. The ref genome was used because...




 calculated and visualized in R v$3.6.1$ [@rcore] using ggplot2 and dplyr v$1.2.1$ [@tidyverse].

# Results

```{r, fig.width = 3, fig.height = 2, fig.align = "left"}
# seq_stat %>%
#   mutate(depth = signif(depth,3),
#          Num.reads = formatC(seq_stat$Num.reads, big.mark = ","),
#          Paired = formatC(seq_stat$Paired, big.mark = ",")) %>%
#   select(ind, depth, Num.reads, Paired) %>%
#   rename(Individual = ind,
#          `Sequencing depth` = depth,
#          `Number of reads` = Num.reads,
#          `Paired reads` = Paired) %>%
  kable("latex", booktabs = T,
        align = c('l', 'c', 'c', 'c'),
        caption = "CAPTION") %>%
  kable_styling(font_size = 10)
```

```{r, fig.width = 5, fig.height = 2, fig.cap = "CAPTION")}



cowplot::plot_grid(p1, p2, nrow = 1, labels = "AUTO", label_size = 10)
```

The mapping rates...

The number of DEGs...

The experimental model ...

Which DEGs...

# Conclusion

The conservation of _Picea rubens_ along the eastern United States...

Future directions for this study must involve...

\pagebreak

# References [^1]

[^1]: This does not exactly follow the MLA citation example you gave, but it was the best `.csl` I could find in time.